\chapter*{Thesis structure}
\addcontentsline{toc}{chapter}{Thesis structure}
\markboth{Thesis structure}{}


The work presented in this thesis is the culmination of five years of research conducted within the bio2Byte group, under the supervision of Wim Vranken. The structure of this thesis reflects the primary focus of my research, organised into four chapters corresponding to six published and submitted articles. The vision of the bio2Byte group has significantly influenced my development as a researcher and, by extension, the contents of this thesis. One key, unifying concept threads through this work: proteins are dynamic entities with no single, static conformation.
% Enjoy the read.
 \vspace{0.8 em}
 % \hrule
% \vspace{0.8 em}
% \\
% \vspace{1cm}

\noindent
\textbf{Chapter \ref{chapter:constava}}:
\\
\textbf{Jose Gavalda-Garcia}, David Bickel, Joel Roca-Martinez, Daniele Raimondi, Gabriele Orlando, and Wim Vranken. Data-driven probabilistic definition of the low energy conformational states of protein residues.
\textit{NAR Genomics and Bioinformatics}, 6(3):lqae082, September 2024. 
\vspace{0.8 em}
\hrule
\vspace{0.8 em}
\noindent
\textbf{Chapter \ref{chapter:b2btools_deployment}}:
\\
\textbf{Jose Gavalda-Garcia}, Adrian Diaz, and Wim Vranken. bio2Byte Tools deployment as a Python package and Galaxy tool to predict protein biophysical properties, July 2024. \textit{arXiv:2405.02136 [q-bio]}.
\\
\\
Luciano Kagami, Joel Roca-Martínez, \textbf{Jose Gavaldá-García}, Pathmanaban Ramasamy, K.~Anton Feenstra, and Wim~F. Vranken. Online biophysical predictions for {SARS}-{CoV}-2 proteins. \textit{BMC Molecular and Cell Biology}, 22(1):23, April 2021.
\\
\\
Luciano Porto Kagami, Gabriele Orlando, Daniele Raimondi, Francois Ancien, Bhawna Dixit, \textbf{Jose Gavaldá-García}, Pathmanaban Ramasamy, Joel Roca-Martínez, Konstantina Tzavella, and Wim Vranken.{b2bTools}: online predictions for protein biophysical features and their conservation.\textit{Nucleic Acids Research}, 49(W1):W52--W59, July 2021.
\vspace{0.8 em}
\hrule
\vspace{0.8 em}
\noindent
\textbf{Chapter \ref{chapter:ambiguous}}:
\\
Joel Roca-Martinez, Tamas Lazar, \textbf{Jose Gavalda-Garcia}, David Bickel, Rita Pancsa, Bhawna Dixit, Konstantina Tzavella, Pathmanaban Ramasamy, Maite Sanchez-Fornaris, Isel Grau, and Wim~F. Vranken.
Challenges in describing the conformation and dynamics of proteins with ambiguous behavior. \textit{Frontiers in Molecular Biosciences}, 9, August 2022.
\vspace{0.8 em}
\hrule
\vspace{0.8 em}
\noindent
\textbf{Chapter \ref{chapter:plddt}}:
\\
\textbf{Jose Gavalda-Garcia}, Bhawna Dixit, Adrián Díaz, An Ghysels, and Wim Vranken. Gradations in protein dynamics captured by experimental NMR are not well represented by AlphaFold2 models and other computational metrics, July 2024. \textit{bioRxiv}, Pages: 2024.07.17.603933 Section: New Results.
