\chapter*{Acknowledgements}
\addcontentsline{toc}{chapter}{Acknowledgements}
\markboth{ACKNOWLEDGEMENTS}{} % Clear the header for this section

Five years come to an end and I finally have the space to look back and reflect on them. These years have been the most challenging of my life, both personally and professionally. The reason why you can now read this thesis is because of those close to me, who have supported me in moments of adversity and celebrated the wins, no matter how big or small. Before I tell you how much I appreciate you and everything you have done for me during these years, I am going to thank myself. I thank myself for not giving up, for becoming kinder to myself, for setting boundaries, for learning that resilience is not enough, and for accepting that reaching out for help is a sign of strength, not weakness. I thank myself for developing in a way that made me survive this experience. 

This PhD would not have been possible without the opportunity given to me by professor Wim Vranken to be part of his group. You have taught me a lot about Belgian culture, beers, and of course wobbling proteins. Thank you for your guidance and support throughout all these years.

Many colleagues have come and gone during these years, and I like to think I have learned something from each and every one of them. Though you have all helped me, some have been especially important in this journey and deserve some extra lines here. Joel, you have been a good friend from the very beginning, and I treasure all the climbs, trips, talks, and drinks that we have shared and I'm happy to see you thrive, now from a distance. Aitor, we have not drank as many beers as we should have --- it's entirely my fault for being a tired boy. I have fond memories of our pre-Covid morning chats whilst you smoked, the advice given, and all the side quests that you suggested for outreach. Thanks to you, Joel and I can say that we have explained our project using a plate of mussels. Paddy, you absolute party beast, you have always put a smile on my face and have taught me, by example, a thing or two about the art of seduction. Please stay true to yourself; you're awesome. David, you rescued me from some of the deepest, scariest moments in my whole PhD, supporting me both emotionally and scientifically. You were my ally when external forces threatened this project and you kept me on track. You've played a transcendent role in my life and helped me recover the spark for doing science and I will forever be grateful for encountering you in the right time and place. Adrian, my \textit{pibe}, you have been one of the biggest reasons to get out of bed every morning. You have been a safe space to express myself when I felt vulnerable and needed to listen to a calming voice. I keep my cow foot \textit{mate} in sight to always have you on my mind. Sophie, my ally in the office, I'm happy to have someone to share opinions with about pretty much anything. Opinionated people are cool and we can scream that together. Roswitha, \textit{il mio pericolo Siciliano}, you were my partner in crime, in and out of work, during my research stay in Valencia and beyond. I hope you're making a list of all the places where we're going to eat and drink when I visit you, I'm already crafting mine for when you come. You have all been fundamental in the completion of this journey --- thank you.

Outside of work, I have made some outstanding friends during my time in Brussels who have made me unplug from work and enjoy life. Justine, Jeroen and Amira, my dear friends, you have given me a blast of a time with dinners, karaoke nights, hikes, birthdays, holidays, drinks, and just being there for me when I needed you. I will forever remember eating leftover dim sum with you, during one of the coldest nights of my life, on a half-sinking boat in the canals of Amsterdam. Also, thank you for not making fun of me for saying that I was finishing my PhD in half a year, continuously for the last 2 years. Llŷr, you have become one of my closest friends and helped me regain control over my physical and mental health by turning me into a runner. All the proof-reading that you have put into this thesis is mental, and your sassy ass comments made me laugh through very tiring times. I'll forever remember that ``Sexy people use Oxford commas''. Jalil, your \textit{Finch} notifications in the morning and all the politically incorrect memes we exchange on Instagram are a driving force beyond my residual serotonin levels --- thank you for all of that. Lucas, you have been a meme dealer, restaurant connoisseur, and, at times, a psychologist. What's better than a friend who takes you to good restaurants and listen to your complaints? Thank you for reassuring that what I felt was valid and natural. Of course, I need to thank Sharkpai, the number one Eurovision Song Contest fan and catalyser of my wildest thoughts, stay savage.

My journey to Brussels started in Amsterdam, where the VU Bioinformatics group adopted me as a student, teacher, and colleague. Again, all of you were critical in my scientific and personal development, but there are some honorary mentions that I'd like to highlight. Juami, you were always ready to turn our office into a musical theatre stage and even managed to make me join you for zumba and modern dance. Your energy is amazing, the parties we organised as founding members of the party committee were unforgettable, and being the paranymph for your PhD defence was one of the happiest, proudest experiences of my life. Anton, your party spirit is unmatchable and only my sofa knows how often we have partied past your last train back home. Your friendship and guidance have continued beyond my stay in Amsterdam. I look forward to joining you on some more runs once my pace matches yours. Sanne, you taught me so much, not only about science but about what make a scientist great. I have kept your words and lessons very present all these years to become a better scientist every day. Halima, probably the only person that likes talking about food as much as I do. You are incredibly smart and well-achieved, yet you have always treated me as an equal and have given me confidence in my skills when I doubted myself. Bas, you're the personification of a giant caring viking. You made me feel like you always had my back, and I'm sure your daughters will feel the most fortunate girls in the world growing up with you as their dad. Working with you was a delight and seeing you in character during DnD sessions was hilarious, please never change. Jaap, you made me tear up when you read your speech for my master's graduation ceremony and even hosted the celebration dinner! You trusted me with responsibilities that made me feel a core part of something important, and you were never too busy to share your knowledge and experiences with me. I still owe you and Hellen a paella when I move back to Amsterdam. 

Amsterdam gave me some of the most important people in my life. Nathalie and Vanessa, I'm so proud of how far you have reached, even when life threw some very objectively strong blows your way. I enjoyed our moments so much, from partying for Koningsdag to running a snowy obstacle race in April, and I can't wait to be near you again. Most importantly, Amsterdam gave me my life partner, Maulo, who joined me in my PhD adventure and has made every moment together magical. You held me when I needed to cry and popped open champagne when we celebrated a win. You have been my rock to hold on to during my most difficult moments and always called out my nonsense when emotions blinded me from reality. You sent me out for a run when I was overwhelmed, cooked for me when I didn't have the energy to nourish myself, and so much more that this book would need a whole new chapter just to capture it all. I love you, I love the memories and the life that we have built together, and I can't wait to spend the rest of our lives together. With Maulo, I also gained an adoptive family: Duangnet, always up for fetching oysters and barbecuing way more food than our stomachs could fit. Willem, reliable and ready to drive through international borders to rescue us from train service outages and bring us home. Sietze-Heng, always sending me Formula 1 memes and tolerating my bullying about anything related to Max Verstappen. You should probably know by now that it's impossible to reason with a chaotic evil Alonso fan.

Before Amsterdam, there was Valencia and my friends made during my bachelor, who are now spread all around the globe. Alba, Lucía, Alejandra, Evy, Miriam and Alejandro, I'm glad that we still find the time to meet online, have our now-traditional Christmas dinners, and remain a part of each other's lives despite the distance. Alba, you also joined me in Amsterdam and we shared a lot of street food at the food stalls in its markets. Lucía, only we could start an iGEM team by mutually reinforcing our random behaviours, epic. Alejandra, you have helped me so much by reassuring me that my thoughts along these years were valid and normal, thank you for being there and putting things into perspective. Evy, you always find a way of making me see things from the other person’s angle. Miriam, you are my inspiration to find healthy ways to unwind after the 9 to 5 grind. Alejandro, you showed me ways to keep my head down and focus on a goal, sharing similar circumstances during the final year of our PhD journey. Meeting you all has turned into a person who likes pineapple on pizza, someone who has an origami Dalek in my childhood bedroom, and one of the few privileged VIPs that could see \textit{La Nit del Foc} from within the safety perimeter year after year during \textit{Fallas}.

Looking back even further, I reach my home town, Nules. A place where I have not officially lived for almost half of my life, after leaving at seventeen. As a kid, you call many people your friends, and the list gets shorter and shorter as you grow up until only the true ones prevail. This selection is even more pronounced when you leave to live hundreds or thousands of kilometres away. Carles, we didn't meet until high school, but that didn't limit how close we would become. I opened my heart to you and your wife Belén during my speech at your wedding, so allow me to update the message to include your beautiful children, Daniel and Pablo. I look forward to seeing them grow up and become as good as you are. Edgar, you also came into my life during high school. Since then, you have been somebody loving, understandable, and reliable with whom I can freely speak my mind and share my insecurities and vulnerabilities. Thank you for always being there for me. Edu, our friendship has survived nearly breaking your collarbone with a cola can so I'd say it's bulletproof (at least cola-proof). I think everybody in this section has built some of their best memories with you, spending our Easter break at your parents' house in the mountain, grilling pigs face and not seeing the light of day because we were playing video games for a week straight. I hope we can build many more of these memories. Diego, an earthquake happens every time you pull out your card to pay for drinks. Let's do that more often, I miss those chaos nights, and I promise I'll try to keep your car clean this time. Pablo, you are my oldest friend and I don't recall having any memories before you. We have blocked roads during student protest and still had time to eat something in Chinatown afterwards. We mutually reinforced our hipster era went to clubs that made us feel as cool as if we had just made it into Berghain. Sam, you're in the Nules gang paragraph too. I feel very fortunate that my language exchange buddy became one of my best friends. My plans on moving to London have been put on hold, but rest assured you're seeing me enough once I have my big boy job and want to see all the plays in the West End, and you better come see me wherever I am too! You all have been with me through it all, believing in me through thick and thin. You have recharged my batteries when they were drained and prepared me to come back and face the challenge that this doctorate has been. Thank you guys, I love you. 

One more look back and I find my family. My uncle Enrique and aunt Alicia, and my cousins Beatriz and Iván. Non-Spanish people can't believe how strongly Spanish families are held together until I tell them that we rarely skip a paella Sunday together. Then, I see my grandfather, Jose María, who valued hard work and intelligence like no other and sponsored every activity that could help my development. Further back and I see my parents, who have given me nothing but unconditional love and support throughout my life. They gave me a childhood with nothing to worry about, so I could focus on being a child and learning. As a young adult, they helped me with my coming of age and to come to terms with myself, with all the love that they had already showed me as a child. They also funded my studies when the grants that I could get were not enough. This trajectory, this thesis and, every other opportunity that I have had have only possible thanks to their hard work, love and generosity. For that, I will be forever grateful. \textit{Gracias papá, gracias mamá, os quiero.}

Finally, I want to commemorate the person to which I dedicate this thesis, my grandmother Brígida, \textit{pati} for her grandchildren because I made up that name for her as a toddler. I came into this world as her first grandchild, just days before my grandfather Enrique passed away. The most important person in her life was gone and there was a crying blob of meat to replace him, me. She raised me and became my second mother, and all her love and kindness permeated into me, shaping core parts of myself. Time passed by, I moved out of home, I moved abroad, her memory started to fade but she kept calling me \textit{``mi campeón''}, ``my champion'', until she couldn't speak any longer. I could take care of her during her last week with us and be at peace that I could repay some of the kindness she gave me growing up. This thesis is dedicated to her because the Jose that completed it would not exist without her. \textit{Te quiere, tu campeón.}